---
title: "Final Project: Exploring Factors Contributing to Positive Changes in Mental Health for Children Transitioning Into Adulthood"
subtitle: "Feature Selection"
author: "Travis Jones, III"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**final_project_feature_selection.Rmd**.

# Feature Selection Introduction
I used the Boruta algorithm for its feature selection.  The problem is that the features aren't ordered by any statistical significance.  For this reason, I decided to use the ordinary least squares model to help both confirm the features selected by the Boruta model, but also to determine the more significant features affecting PositiveChange.  I'm hoping to select the common features between the two models, and use the ordering information from the OLS model.  

I've decided against using the RFE model because Boruta is based on the RFE model.  

```{r Common, warning=F, message=F, include=F, child='final_project_common.Rmd'}
```

```{r Header, warning=F, message=F}
library(Boruta)
library(mlbench)

# return the OLS stepwise model given the dataframe
run_stepwise <- function(df_run) {
  set.seed(1234)
  base.mod <- lm(PositiveChange ~ 1, data = df_run)
  # Define the full model - including all predictors
  all.mod <- lm(PositiveChange ~ ., data = df_run)
  ols_step <- lm(PositiveChange ~ ., data = df_run)
  ols_step <- step(base.mod, scope = list(lower = base.mod, upper = all.mod), direction = 'both', k = 2, trace = F)
}

# translate the OLS model formula from obscure fieldnames into their more descriptive descriptions
formula_to_vector_descriptions <- function(formula) {
  vector_untranslated <- trimws(strsplit(as.character(formula[3]), "\\+")[[1]])
  vector_translation <- lapply(vector_untranslated, fieldname_to_description)
  return(vector_translation)
}

# translate the OLS model summary using the descriptive descriptions instead of their obscure fieldnames
# sort by p-values descending and add stars based upon statistical significance
# p-value < 0.001   -> ***
# p-value < 0.01    -> **
# p-value < 0.05    -> *
# p-value < 0.1     -> .
# else              -> -
lm_to_summary_model <- function(lmodel) {
  sum_ols <- summary(lmodel)
  df_ols_model <- as.data.frame(as.matrix(lapply(names(lmodel$coefficients), fieldname_to_description)))
  vect_pvals <- unname(sum_ols$coefficients[, 4])
  vect_stars <- lapply(vect_pvals, function(x) {
      if(x > 0 & x < 0.001)
        return("***")
      else if(x >= 0.001 & x < 0.01)
        return("**")
      else if(x >= 0.01 & x < 0.05)
        return("*")
      else if(x >= 0.05 & x < 0.1)
        return(".")
      else if(x >= 0.1 & x < 1)
        return("_")
    })
  df_ols_model <- cbind(Variables = df_ols_model$V1, pvalues = vect_pvals, Stars = vect_stars)
  as.data.frame(df_ols_model[order(vect_pvals), ])
}

# crossvalidate the OLS model
crossvalidate_ols <- function(df_train, df_test) {
  yTest <- df_test$PositiveChange
  XTest <- df_test  [ , !names(df_test) %in% "PositiveChange"]
  ols_validate <- run_stepwise(df_train)
  betaHatOLS_step <- ols_validate$coefficients
  var_step <- colnames(ols_validate$model)[-1]
  XTestOLS_step <- cbind(rep(1, nrow(XTest)), XTest[,var_step])
  predOLS_step <- as.matrix(XTestOLS_step)%*%as.matrix(betaHatOLS_step)
  testMSEOLS_step <- mean((predOLS_step - yTest)^2)
  # Report MSE OLS Stepwise feature selection
  testMSEOLS_step
  pred2 <- predict(ols_validate, as.data.frame(XTest))
  any(pred2 == predOLS_step)
}

# TODO: crossvalidate the Boruta model
crossvalidate_boruta <- function(df_train, df_test) {
  yTest <- df_test$PositiveChange
  XTest <- df_test  [ , !names(df_test) %in% "PositiveChange"]
  boruta_validate <- Boruta(PositiveChange ~ ., df_train)
}
```

## Boruta - Full Dataset
```{r BorutaFull, warning=F, message=F}
set.seed(1234)
boruta_full <- Boruta(PositiveChange ~ ., data = df_full_scaled)
print(boruta_full)
#boruta_full <- TentativeRoughFix(boruta_full)
descriptions_full <- formula_to_vector_descriptions(getConfirmedFormula(boruta_full))
paste(descriptions_full, collapse = " + ")
#print(boruta_full$finalDecision[boruta_full$finalDecision %in% c("Confirmed", "Tentative")])

df_full_long <- gather(as.data.frame(boruta_full$ImpHistory), feature, measurement)
feature_full <- df_full_long$feature
feature_label_full <- unlist(lapply(feature_full, fieldname_to_description))
measurement_full <- unlist(lapply(df_full_long$measurement, function(x) { if(x == "-Inf") x = 0; return(x)}))
df_full_plotly <- data.frame(Feature = feature_full, FeatureLabel = feature_label_full, Measurement = measurement_full)
plot_ly(df_full_plotly, x = ~FeatureLabel, y = ~Measurement, color = ~Feature, marker = list(size = 2), type = "box") %>%
  layout(title="Box-and-whisker Plots Across All Features",
           xaxis = list(title="Features"),
           yaxis = list(title="Importance"),
           showlegend=F)
```

## Boruta - Objective Dataset
```{r BorutaObjective, warning=F, message=F}
set.seed(1234)
boruta_objective <- Boruta(PositiveChange ~ ., data = df_objective_scaled)
print(boruta_objective)
#boruta_objective <- TentativeRoughFix(boruta_objective)
descriptions_objective <- formula_to_vector_descriptions(getConfirmedFormula(boruta_objective))
paste(descriptions_objective, collapse = " + ")
#print(boruta_objective$finalDecision[boruta_objective$finalDecision %in% c("Confirmed", "Tentative")])

df_objective_long <- gather(as.data.frame(boruta_objective$ImpHistory), feature, measurement)
feature_objective <- df_objective_long$feature
feature_label_objective <- unlist(lapply(feature_objective, fieldname_to_description))
measurement_objective <- unlist(lapply(df_objective_long$measurement, function(x) { if(x == "-Inf") x = 0; return(x)}))
df_objective_plotly <- data.frame(Feature = feature_objective, FeatureLabel = feature_label_objective, Measurement = measurement_objective)
plot_ly(df_objective_plotly, x = ~FeatureLabel, y = ~Measurement, color = ~Feature, marker = list(size = 2), type = "box") %>%
  layout(title="Box-and-whisker Plots Across Objective Features",
           xaxis = list(title="Features"),
           yaxis = list(title="Importance"),
           showlegend=F)
```

## Stepwise Feature Selection - Full Dataset

```{r StepwiseFull, warning=F, message=F}
ols_full <- run_stepwise(df_full_scaled)
df_ols_model_full_trans <- lm_to_summary_model(ols_full)
# get the shortlisted variable
# ols_confirmed_vars_full <- names(unlist(ols_full[[1]]))
# # remove the intercept
# ols_confirmed_vars_full <- ols_confirmed_vars_full[!ols_confirmed_vars_full %in% "(Intercept)"]
# ols_confirmed_vars_full <- unlist(lapply(ols_confirmed_vars_full, fieldname_to_description))
# print(ols_confirmed_vars_full)
# estimate variable importance
varImp_ols_full <- varImp(ols_full, scale=FALSE)
rownames(varImp_ols_full) <- unlist(lapply(rownames(varImp_ols_full), fieldname_to_description))
# summarize importance
print(varImp_ols_full)
# crossvalidate between the training and testing data
crossvalidate_ols(df_full_train_scaled, df_full_test_scaled)
```

## Stepwise Feature Selection - Objective Dataset
```{r StepwiseObjective, warning=F, message=F}
ols_objective <- run_stepwise(df_objective_scaled)
df_ols_model_objective_trans <- lm_to_summary_model(ols_objective)
# get the shortlisted variable
# ols_confirmed_vars_objective <- names(unlist(ols_objective[[1]]))
# # remove the intercept
# ols_confirmed_vars_objective <- ols_confirmed_vars_objective[!ols_confirmed_vars_objective %in% "(Intercept)"]
# ols_confirmed_vars_objective <- unlist(lapply(ols_confirmed_vars_objective, fieldname_to_description))
# print(ols_confirmed_vars_objective)
# estimate variable importance
varImp_ols_objective <- varImp(ols_objective, scale=FALSE)
rownames(varImp_ols_objective) <- unlist(lapply(rownames(varImp_ols_objective), fieldname_to_description))
# summarize importance
print(varImp_ols_objective)
# crossvalidate between the training and testing data
crossvalidate_ols(df_objective_train_scaled, df_objective_test_scaled)
```

# Summary
As you will see below, the full featured dataset contains more data than the "objective" counterparts.  Each dataset provides a different view into the question at hand.
```{r Summary, warning=F, message=F}
# combine the two common features for full dataset
df_full_feauture_selection_combined <- as.data.frame(df_ols_model_full_trans[is.element(set = df_ols_model_full_trans$Variables, el = as.matrix(descriptions_full)), ])

# combine the two common features for objective dataset
df_objective_feature_selection_combined <- as.matrix(df_ols_model_objective_trans[is.element(set = df_ols_model_objective_trans$Variables, el = as.matrix(descriptions_objective)), ])

print("Combination of Boruta and OLS Full Datasets ")
df_full_feauture_selection_combined
print("Combination of Boruta and OLS Objective Datasets ")
df_objective_feature_selection_combined
```
