---
title: "Final Project"
subtitle: "<h2><u>Data Science and Predictive Analytics (HS650), Fall 2021</u></h2>"
author: "<h3>Kevin Wu</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**Final Project.Rmd**.

 * HW #3
 * Fall 2021, DSPA (HS650)
 * Name: Kevin Wu
 * SID: ####0012 (last 4 digits only)
 * UMich E-mail: kevinkwu@umich.edu
 * I certify that the following paper represents my own independent work and conforms with the guidelines of academic honesty described in the UMich student handbook.
 * Remember that students are allowed and encouraged to discuss, on a conceptual level, the problems with your class mates, however, this can not involve the exchange of actual code, printouts, solutions, e-mails or other explicit electronic or paper handouts.


# Final Project Introduction
dependent variable general sense of well being
no quantitative way to measure that factor nor clinical predictors
survey post ACA that measures this in Michigan
unfortunately nonthing prior to establish baseline
identify predictive indicators of well being by studying those with higher levels vs those with lower levels
inherent geo-socio-economic biases 

```{r Introduction, warning=F, message=F}


#---------------------------------------------------------------------------
# factor_to_numeric_icpsr.R
# 2012/12/06
#
# Convert R factor variable back to numeric in an ICPSR-produced R data
# frame. This works because the original numeric codes were prepended by
# ICSPR to the factor levels in the process of converting the original
# numeric categorical variable to factor during R data frame generation.
#
# REQUIRES add.value.labels function from prettyR package
#    http://cran.r-project.org/web/packages/prettyR/index.html
#
#
# Substitute the actual variable and data frame names for da99999.0001$MYVAR
# placeholders in syntax below.
#
#    data frame = da99999.0001
#    variable   = MYVAR
#
#
# Line-by-line comments:
#
# (1) Load prettyR package
#
# (2) Create object (lbls) containing the factor levels for the specified
#     variable.  Sort will be numeric as original codes (zero-padded, if
#     necessary) were preserved in the factor levels.
#
# (3) Strip original codes from lbls, leaving only the value labels, e.g.,
#       "(01) STRONGLY DISAGREE" becomes "STRONGLY DISAGREE"
#
# (4) Strip labels from data, leaving only the original codes, e.g.,
#       "(01) STRONGLY DISAGREE" becomes "1"
#
#     Then, coerce variable to numeric
#
# (5) Add value labels, making this a named numeric vector
#---------------------------------------------------------------------------
#lbls <- sort(levels(da99999.0001$MYVAR))
#lbls <- (sub("^\\([0-9]+\\) +(.+$)", "\\1", lbls))
#da99999.0001$MYVAR <- as.numeric(sub("^\\(0*([0-9]+)\\).+$", "\\1", da99999.0001$MYVAR))
#da99999.0001$MYVAR <- add.value.labels(da99999.0001$MYVAR, lbls)

library(tm)
library(stringi)
library(prettyR)
library(plotly)
load(file = ".RData")

```

# Data Preparation
```{r Data Preparation, warning=F, message=F}
#load RDA data frames into environment
con2019Q1 <- url("http://kevinwurn.github.io/UMich-HS650/final_project_files/2019Q1HealthReformMonitoringSurvey-da37922.0001.rda")
load(con2019Q1)
close(con2019Q1)
con2019Q3 <- url("http://kevinwurn.github.io/UMich-HS650/final_project_files/2019Q3HealthReformMonitoringSurvey-da37630.0001.rda")
load(con2019Q3)
close(con2019Q3)

#rename dataframes
df2019Q1 <- da37922.0001
df2019Q3 <- da37630.0001

#add year/quarter stamps - change RID 
df2019Q1$year <- 2019
df2019Q1$quarter <- 0
df2019Q1$quarter <- as.factor(df2019Q1$quarter)
df2019Q1$RID <- paste0("A", df2019Q1$RID)

df2019Q3$year <- 2019
df2019Q3$quarter <- 1
df2019Q3$quarter <- as.factor(df2019Q3$quarter)
df2019Q3$RID <- paste0("B", df2019Q3$RID)

#merge the two dataframes
colnames2019Q1 <- names(df2019Q1)
colnames2019Q3 <- names(df2019Q3)

#
removenames2019Q1 <- removeWords(colnames2019Q1,colnames2019Q3)
colnames2019Q1 <- removeWords(colnames2019Q1, removenames2019Q1)

removenames2019Q3 <- removeWords(colnames2019Q3,colnames2019Q1)
colnames2019Q3 <- removeWords(colnames2019Q3, removenames2019Q3)

colnamesInnerJoin <- stri_remove_empty_na(colnames2019Q3)

df2019Q1 <- df2019Q1[colnamesInnerJoin]
df2019Q3 <- df2019Q3[colnamesInnerJoin]
df2019 <- rbind(df2019Q1, df2019Q3)

#clean up / factorialize the data1
for (i in 1:ncol(df2019)) {
  column <- df2019[,i]
  df2019[,i] <- as.numeric(sub("^\\(0*([0-9]+)\\).+$", "\\1", column))
}
```

# Data Summary
median is 0 days.. good.  50% respondents.

CopyQ3

CopyQ4

histogram shows... increments of 5... why?  because.. people naturally think in terms of 5 days.
interestingly, despite it's right skewness... mean is still   4-5 days out of 30 days - roughly 15% which is still telling... that when people are sick, they are sick.

Doesn't show it, but

```{r Data Summary, message=FALSE, warning=FALSE}
#had to trim out the NA to calculate the mean/median, otherwise, I would receive NA as the result.
mean(df2019$Q3, na.rm = T)
median(df2019$Q3, na.rm = T)
length(df2019$Q3)


plot_ly() %>%
  add_histogram(df2019$Q3) %>%
  layout(title="Histogram over 30 Days Where Mental Health Was Not Good",
    xaxis=list(title="Number of Days"),
    yaxis=list(title="Number of Participants")
  )

#HIST1 - A Place to go for healthcare advice
dfSubPlaceToGo <- subset(df2019, Q4 == 1 || Q4 == 2)
#HIST2 - No place to go for healthcare advice
dfSubNoPlaceToGo <- subset(df2019, Q4 == 3)
plot_ly() %>%
  #HIST1
  add_histogram(~dfSubPlaceToGo$Q3, name="Those with Access to Advice") %>%
  #HIST2
  add_histogram(~dfSubNoPlaceToGo$Q3, name="Those without Access") %>%
  layout(title="Histogram over 30 Days Where Mental Health Was Not Good",
    xaxis=list(title="Number of Days"),
    yaxis=list(title="Number of Participants")
  )

nullcount <- 0
for (i in 1:ncol(df2019)) {
  column <- df2019[,i]
  columnCount <- sum(is.na(column))
  if (columnCount > 0)
    nullcount <- nullcount + 1
  print(paste0(names(df2019[,i]), " has ", columnCount, " null entries"))
}
print (nullcount)
