---
title: "Final Project: Exploring Factors Contributing to Positive Changes in Mental Health for Children Transitioning Into Adulthood"
subtitle: "Trees"
author: "Kevin Wu"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**final_project_trees.Rmd**.


# Classification Trees
```{r Introduction, warning=F, message=F, include=F, child='final_project_common.Rmd'}
```

## Helper Functions and Libraries
```{r Header, warning=F, message=F}
library(C50)
library(rpart.plot)
library(rattle)
library(neuralnet)
library(kernlab)

# colnames(df_full_train) <- lapply(colnames(df_full_train[, -267]), fieldname_to_description)
# colnames(df_full_test) <- lapply(colnames(df_full_test[, -267]), fieldname_to_description)

df_full_train_new <- df_full_train[ , !names(df_full_train) %in% "PositiveChange"]
df_full_test_new <- df_full_test[ , !names(df_full_test) %in% "PositiveChange"]
df_objective_train_new <- df_objective_train[ , !names(df_objective_train) %in% "PositiveChange"]
df_objective_test_new <- df_objective_test[ , !names(df_objective_test) %in% "PositiveChange"]

confusionMatrix_to_vector <- function(mat) {
  return(c(
    round(mat$overall["Accuracy"], digits = 4), 
    round(mat$overall["Kappa"], digits = 4), 
    round(mat$byClass["Sensitivity"], digits = 4), 
    round(mat$byClass["Specificity"], digits = 4)
  ))
}

highlight_column <- function(col) {
  cell_spec(col, background = ifelse(col == max(col), "yellow", "white"))
}

```
## Neural Net - Full Dataset
```{r NNFull, warning=F, message=F}
net_full <- neuralnet(PositiveChange ~ ., df_full_train, hidden=10, threshold=0.1)
plot(net_full, rep="best")
pred_net_full <- compute(net_full, df_full_test)
# tbl_net_full <- table(pred_net_full$net.result, df_full_test$PositiveChange)
cor(pred_net_full$net.result, as.numeric(df_full_test$PositiveChange))
```

## Neural Net - Objective Dataset
```{r NNObjective, warning=F, message=F}
net_objective <- neuralnet(PositiveChange ~ ., df_objective_train, hidden=5, threshold=0.1)
plot(net_objective, rep="best")
pred_net_objective <- compute(net_objective, df_objective_test)
cor(pred_net_objective$net.result, as.numeric(df_objective_test$PositiveChange))
```

## C5.0 - Full Dataset
```{r C50Full, warning=F, message=F}
c50_model_full <- C5.0(df_full_train_new, df_full_train$PositiveChange)
c50_model_full
summary(c50_model_full)
pred_C50_full <- predict(c50_model_full, df_full_test_new) 
tbl_C50_full <- table(pred_C50_full, df_full_test$PositiveChange)
conf_C50_full <- confusionMatrix(tbl_C50_full)
conf_C50_full
```

## C5.0 - Objetive Dataset
```{r C50Objective, warning=F, message=F}
c50_model_objective <- C5.0(df_objective_train_new, df_objective_train$PositiveChange)
c50_model_objective
summary(c50_model_objective)
pred_C50_objective <- predict(c50_model_objective, df_objective_test_new) 
tbl_C50_objective <- table(pred_C50_objective, df_objective_test$PositiveChange)
conf_C50_objective <- confusionMatrix(tbl_C50_objective)
conf_C50_objective
```

## RPart - Full Dataset
```{r RPartFull, warning=F, message=F}
rpart_model_full <- rpart(PositiveChange ~ ., data = df_full_train, cp=0.01) 
rpart_model_full
fancyRpartPlot(rpart_model_full, cex = 1, caption = "Fancy Rpart Plot (Full Dataset)")
pred_rpart_full <- predict(rpart_model_full, df_full_test, type = 'class')
tbl_rpart_full <- table(pred_rpart_full, df_full_test$PositiveChange)
conf_rpart_full <- confusionMatrix(tbl_rpart_full)
conf_rpart_full

rpart_model_pruned_full <- prune(rpart_model_full, cp = rpart_model_full$cptable[which.min(rpart_model_full$cptable[,"xerror"]),"CP"])
fancyRpartPlot(rpart_model_pruned_full, cex = 1, caption = "Fancy Rpart Plot (Full Dataset) Pruned")
pred_rpart_pruned_full <- predict(rpart_model_pruned_full, df_full_test, type = 'class')
tbl_rpart_pruned_full <- table(pred_rpart_pruned_full, df_full_test$PositiveChange)
conf_rpart_pruned_full <- confusionMatrix(tbl_rpart_pruned_full)
conf_rpart_pruned_full
```

## RPart - Objective Dataset
```{r RPartObjective, warning=F, message=F}
rpart_model_objective <- rpart(PositiveChange ~ ., data = df_objective_train, cp=0.01) 
rpart_model_objective
fancyRpartPlot(rpart_model_objective, cex = 1, caption = "Fancy Rpart Plot (Objective Dataset)")
pred_rpart_objective <- predict(rpart_model_objective, df_objective_test, type = 'class')
tbl_rpart_objective <- table(pred_rpart_objective, df_objective_test$PositiveChange)
conf_rpart_objective <- confusionMatrix(tbl_rpart_objective)
conf_rpart_objective

rpart_model_pruned_objective <- prune(rpart_model_objective, cp = rpart_model_objective$cptable[which.min(rpart_model_objective$cptable[,"xerror"]),"CP"])
fancyRpartPlot(rpart_model_pruned_objective, cex = 1, caption = "Fancy Rpart Plot (Objective Dataset) Pruned")
pred_rpart_pruned_objective <- predict(rpart_model_pruned_objective, df_objective_test, type = 'class')
tbl_rpart_pruned_objective <- table(pred_rpart_pruned_objective, df_objective_test$PositiveChange)
conf_rpart_pruned_objective <- confusionMatrix(tbl_rpart_pruned_objective)
conf_rpart_pruned_objective
```

## RBF - Full Dataset
```{r RBFFull, warning=F, message=F}
rbf_full <- ksvm(PositiveChange ~ ., data = df_full_train, kernel = "rbfdot")
pred_rbf_full <- predict(rbf_full, df_full_test)
tbl_rbf_full <- table(pred_rbf_full, df_full_test$PositiveChange)
conf_rbf_full <- confusionMatrix(tbl_rbf_full)
conf_rbf_full
```

## RBF - Objective Dataset
```{r RBFObjective, warning=F, message=F}
rbf_objective <- ksvm(PositiveChange ~ ., data = df_objective_train, kernel = "rbfdot")
pred_rbf_objective <- predict(rbf_objective, df_objective_test)
tbl_rbf_objective <- table(pred_rbf_objective, df_objective_test$PositiveChange)
conf_rbf_objective <- confusionMatrix(tbl_rbf_objective)
conf_rbf_objective
```

# Classification Trees Summary
As shown below, the C5.0 model performed the best for the full dataset.  However, for the objective dataset, RBF performed better with the C5.0 performance dropping quite significantly.  Overall the models performed much better with the full dataset.  For some reason, the Neural Net performed quite poorly.
```{r Summary, warning=F, message=F}
vect_C50_full <- confusionMatrix_to_vector(conf_C50_full)
vect_rpart_full <- confusionMatrix_to_vector(conf_rpart_full)
vect_rpart_pruned_full <- confusionMatrix_to_vector(conf_rpart_pruned_full)
vect_rbf_full <- confusionMatrix_to_vector(conf_rbf_full)
df_stats_full <- data.frame(C50 = vect_C50_full, RPart = vect_rpart_full, RPartPruned = vect_rpart_pruned_full, RBF = vect_rbf_full)
df_stats_full <- as.data.frame(t(df_stats_full))
df_stats_full$Accuracy <- highlight_column(df_stats_full$Accuracy)
df_stats_full$Kappa <- highlight_column(df_stats_full$Kappa)
df_stats_full$Sensitivity <- highlight_column(df_stats_full$Sensitivity)
df_stats_full$Specificity <- highlight_column(df_stats_full$Specificity)
df_stats_full %>%
  kbl(escape = F) %>%
  kable_styling()

vect_C50_objective <- confusionMatrix_to_vector(conf_C50_objective)
vect_rpart_objective <- confusionMatrix_to_vector(conf_rpart_objective)
vect_rpart_pruned_objective <- confusionMatrix_to_vector(conf_rpart_pruned_objective)
vect_rbf_objective <- confusionMatrix_to_vector(conf_rbf_objective)
df_stats_objective <- data.frame(C50 = vect_C50_objective, RPart = vect_rpart_objective, RPartPruned = vect_rpart_pruned_objective, RBF = vect_rbf_objective)
df_stats_objective <- as.data.frame(t(df_stats_objective))
df_stats_objective$Accuracy <- highlight_column(df_stats_objective$Accuracy)
df_stats_objective$Kappa <- highlight_column(df_stats_objective$Kappa)
df_stats_objective$Sensitivity <- highlight_column(df_stats_objective$Sensitivity)
df_stats_objective$Specificity <- highlight_column(df_stats_objective$Specificity)
df_stats_objective %>%
  kbl(escape = F) %>%
  kable_styling()
```