---
title: "Final Project"
subtitle: "Clustering"
author: "Kevin Wu"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**final_project_clustering.Rmd**.

# Clustering Nearest Neighbor
<p>I hope to use clustering to determine the combination of features that contribute to PositiveChange.  However, because clustering deals primarily with interval variables (because of distances that need to be calculated), it is impossible to calculate distances based upon nominal variables with arbitrary values.  For this reason, I have separated out the interval variables from my dataset, and one-hot encoded the nominal variables in the dataset, and combined them.</p>
<p>With this dataset, I trained and predicted the k using the kNN model starting at the default k value.  Using the optimal k number of clusters to use, I used this new k value to determine the exact features that had determined the PositiveChange.  Using the k-means model and the optimal k-value, the cluster centers became the basis of my analysis.  Calculating the standard deviation of the cluster centers, I was able to categorize these features into the following separate buckets: strong inverse associations, inverse associations, associations, and strong associations.

```{r Common, warning=F, message=F, include=F, child='final_project_common.Rmd'}
```
## Helper Functions and Libraries
```{r Header, warning=F, message=F}
# kNN libraries
library(class)
library(gmodels)
library(e1071)

# k-means libraries
library(cluster)
library(matrixStats)


# k-means functions
kpp_init <- function(dat, K) {
  x = as.matrix(dat)
  n = nrow(x)
  # Randomly choose a first center
  centers = matrix(NA, nrow=K, ncol=ncol(x))
  set.seed(1234)
  centers[1,] = as.matrix(x[sample(1:n, 1),])
  for (k in 2:K) {
    # Calculate dist^2 to closest center for each point
    dists = matrix(NA, nrow=n, ncol=k-1)
    for (j in 1:(k-1)) {
      temp = sweep(x, 2, centers[j,], '-')
      dists[,j] = rowSums(temp^2)
    }
    dists = rowMins(dists)
    # Draw next center with probability proportional to dist^2
    cumdists = cumsum(dists)
    prop = runif(1, min=0, max=cumdists[n])
    centers[k,] = as.matrix(x[min(which(cumdists > prop)),])
  }
  return(centers)
}

plot_silhouette <- function(df, K, bKPP = T) {
  set.seed(1234)
  print(K)
  if(bKPP)
    clusters <- kmeans(df, kpp_init(df, K), iter.max=10, algorithm='Lloyd')
  else
    clusters <- kmeans(df, K)
  dis <- dist(df)
  sil <- silhouette(clusters$cluster, dis)
  summary(sil)
  fviz <- factoextra::fviz_silhouette(sil, label=T, palette = "jco", ggtheme = theme_classic())
  result <- str_split(fviz$labels$title, pattern = " ")[[1]][8]
  returned <- list(fviz, clusters, result)
  return(returned)
}

plot_tuning <- function(df, K, bKPP = T) {
  mat = matrix(0, nrow = K)
  for(i in 2:K) {
    set.seed(1234)
    if(bKPP)
      clust_kpp = kmeans(df, kpp_init(df, i), iter.max=10, algorithm='Lloyd')
    else
      clust_kpp = kmeans(df, i)
    dis <- dist(df)
    sil = silhouette(clust_kpp$cluster, dis)
    mat[i] = mean(as.matrix(sil)[,3])
  }
  colnames(mat) <- c("Avg_Silhouette_Value")
  df <- data.frame(k = 2:K, sil = mat[2:K])
  plot_ly(df, x = ~k, y = ~sil, type = 'scatter', mode = 'lines', name='Silhouette') %>%
    layout(title="Average Silhouette Graph")
}

explicate_clusters <- function(clusters) {
  df <- as.data.frame(t(clusters$centers))
  rowNames <- rownames(df)
  cluster_num <- length(clusters$size)
  colnames(df) <- paste0("Cluster",c(1:cluster_num))
  p <- plot_ly()
  for(i in 1:ncol(df)) {
    p <- p %>% add_trace(data = df, x = rownames(df), y = ~df[ , i], type = "bar", name = paste0("Cluster", i))
  }
  p <- p %>% layout(title="Explicating Derived Cluster Labels", yaxis = list(title = 'Cluster Centers'), barmode = 'group')
  p
}

matrix_centers_cleaned_up <- function(matrix_centers) {
  new_centers <- matrix_centers
  for(c in 1:ncol(matrix_centers)) {
    name <- colnames(matrix_centers)[c]
    vect_name <- str_split(name, "\\.", n = 2)
    description <- fieldname_to_description(vect_name[[1]][1])
    value_description <- value_fieldname_to_description(vect_name[[1]][2], vect_name[[1]][1])
    colnames(new_centers)[c] <- paste0(description, " - ", value_description)
  }
  return(new_centers)
}

print_centers_summary <- function(matrix_centers) {
  s <- sd(matrix_centers)
  num_clusters <- nrow(matrix_centers)
  vect_clusters <- vector(mode="character", length = num_clusters)
  vect_strong_inverse_association <- vector(mode="character", length = num_clusters)
  vect_inverse_association <- vector(mode="character", length = num_clusters)
  vect_association <- vector(mode="character", length = num_clusters)
  vect_strong_association <- vector(mode="character", length = num_clusters)
  
  for(i in 1:ncol(matrix_centers)) {
    for(j in 1:num_clusters) {
      vect_clusters[j] <- paste0("Cluster ", j)
      if(matrix_centers[j,i] <= -2 * s)
        if(vect_strong_inverse_association[j] == "")
          vect_strong_inverse_association[j] = colnames(matrix_centers)[i]
        else
          vect_strong_inverse_association[j] <- paste0(vect_strong_inverse_association[j], " + ", colnames(matrix_centers)[i])
      else if(matrix_centers[j,i] > -2 * s && matrix_centers[j,i] <= -s)
        if(vect_inverse_association[j] == "")
          vect_inverse_association[j] = colnames(matrix_centers)[i]
        else
          vect_inverse_association[j] <- paste0(vect_inverse_association[j], " + ", colnames(matrix_centers)[i])
      else if(matrix_centers[j,i] >= s && matrix_centers[j,i] < 2 *s)
        if(vect_association[j] == "")
          vect_association[j] = colnames(matrix_centers)[i]
        else
          vect_association[j] <- paste0(vect_association[j], " + ", colnames(matrix_centers)[i])
      else if (matrix_centers[j,i] >= 2 * s)
        if(vect_strong_association[j] == "")
          vect_strong_association[j] = colnames(matrix_centers)[i]
        else
          vect_strong_association[j] <- paste0(vect_strong_association)
      
    }
  }
  df_printout <- data.frame(Cluster = vect_clusters, StrongInverseAssociation = vect_strong_inverse_association, InverseAssociation = vect_inverse_association, Association = vect_association, StrongAssociation = vect_strong_association)
  kable(df_printout)
}

print_sensitivity_specificity <- function(ct, K) {
  ct_TN <- ct$prop.row[1, 1]
  ct_FP <- ct$prop.row[1, 2]
  ct_FN <- ct$prop.row[2, 1]
  ct_TP <- ct$prop.row[2, 2]
  ct_sensi <- ct_TN/(ct_TN+ct_FP) 
  ct_speci <- ct_TP/(ct_TP+ct_FN)
  print(paste0("kNN model k=", K, " Sensitivity=", ct_sensi))
  print(paste0("kNN model k=", K, " Specificity=", ct_speci))  
}


```

## kNN - Full Dataset
```{r kNNFull, warning=F, message=F}
labels_train_full <- df_full_train_onehot[ , "PositiveChange.Positive.Change"]
labels_test_full <- df_full_test_onehot[ , "PositiveChange.Positive.Change"]
df_full_train_onehot_new <- df_full_train_onehot[ , !names(df_full_train_onehot) %in% "PositiveChange.Positive.Change"]
df_full_test_onehot_new <- df_full_test_onehot[ , !names(df_full_test_onehot) %in% "PositiveChange.Positive.Change"]

new_k_full <- k_full_onehot
pred_full <- knn(train = df_full_train_onehot_new, test = df_full_test_onehot_new, cl = labels_train_full, k = new_k_full)
ct_full <- CrossTable(x = labels_test_full, y = pred_full, prop.chisq = F)
print_sensitivity_specificity(ct_full, new_k_full)

tuning_full = tune.knn(x = df_full_train_onehot_new, y = as.factor(labels_train_full), k = 1:k_full_onehot)
tuning_full
new_k_full <- tuning_full$best.model$k
pred_full <- knn(train = df_full_train_onehot_new, test = df_full_test_onehot_new, cl = labels_train_full, k = new_k_full)
ct_full <- CrossTable(x = labels_test_full, y = pred_full, prop.chisq = F)
print_sensitivity_specificity(ct_full, new_k_full)
```

## kNN - Objective Dataset
```{r kNNObjective, warning=F, message=F}
labels_train_objective <- df_objective_train_onehot[ , "PositiveChange.Positive.Change"]
labels_test_objective <- df_objective_test_onehot[ , "PositiveChange.Positive.Change"]
df_objective_train_onehot_new <- df_objective_train_onehot[ , !names(df_objective_train_onehot) %in% "PositiveChange.Positive.Change"]
df_objective_test_onehot_new <- df_objective_test_onehot[ , !names(df_objective_test_onehot) %in% "PositiveChange.Positive.Change"]

new_k_objective <- k_objective_onehot
pred_objective <- knn(train = df_objective_train_onehot_new, test = df_objective_test_onehot_new, cl = labels_train_objective, k = new_k_objective)
ct_objective <- CrossTable(x = labels_test_objective, y = pred_objective, prop.chisq = F)
print_sensitivity_specificity(ct_objective, new_k_objective)
tuning_objective = tune.knn(x = df_objective_train_onehot_new, y = as.factor(labels_train_objective), k = 1:new_k_objective)
tuning_objective
new_k_objective <- tuning_objective$best.model$k
pred_objective <- knn(train = df_objective_train_onehot_new, test = df_objective_test_onehot_new, cl = labels_train_objective, k = new_k_objective)
ct_objective <- CrossTable(x = labels_test_objective, y = pred_objective, prop.chisq = F)
print_sensitivity_specificity(ct_objective, new_k_objective)
```


## k-Means - Full Dataset
```{r kMeansFull, warning=F, message=F}
# new_k_full <- 4
if(new_k_full < 2) new_k_full <- 4
returned_list_F_full <- plot_silhouette(df_full_onehot, K=new_k_full, bKPP=F)
returned_list_F_full[[1]]
returned_list_T_full <- plot_silhouette(df_full_onehot, K=new_k_full, bKPP=T)
returned_list_T_full[[1]]
ifelse(returned_list_T_full[[3]] >= returned_list_F_full[[3]], bKPP_full <- T, bKPP_full <- F)
plot_tuning(df_full_onehot, K=k_full_onehot, bKPP=bKPP_full)
returned_list_full <- plot_silhouette(df_full_onehot, K=new_k_full, bKPP=bKPP_full)
returned_list_full[[3]]
returned_list_full[[1]]
clusters_full <- returned_list_full[[2]]
new_centers_full <- matrix_centers_cleaned_up(clusters_full$centers)
new_centers_full
# StrongInverseAssociation <= -2s
# InverseAssociation > -2s && <= -s
# Association >= s && < 2s
# StrongAssociation >= 2s
print_centers_summary(new_centers_full)

# explicate_clusters(clusters_full)
```

## k-Means - Objective Dataset
```{r kMeansObjective, warning=F, message=F}
# new_k_objective <- 3
if(new_k_objective < 2) new_k_objective <- 3
returned_list_F_objective <- plot_silhouette(df_objective_onehot, K=new_k_objective, bKPP=F)
returned_list_F_objective[[1]]
returned_list_T_objective <- plot_silhouette(df_objective_onehot, K=new_k_objective, bKPP=T)
returned_list_T_objective[[1]]
ifelse(returned_list_T_objective[[3]] >= returned_list_F_objective[[3]], bKPP_objective <- T, bKPP_objective <- F)
plot_tuning(df_objective_onehot, K=k_objective_onehot, bKPP=bKPP_objective)
returned_list_objective <- plot_silhouette(df_objective_onehot, K=new_k_objective, bKPP=bKPP_objective)
returned_list_objective[[3]]
returned_list_objective[[1]]
clusters_objective <- returned_list_objective[[2]]
new_centers_objective <- matrix_centers_cleaned_up(clusters_objective$centers)
new_centers_objective
# StrongInverseAssociation <= -2s
# InverseAssociation > -2s && <= -s
# Association >= s && < 2s
# StrongAssociation >= 2s
print_centers_summary(new_centers_objective)

# explicate_clusters(clusters_objective)
```



# Summary
<p>There are numerous features, in the various buckets and are difficult to parse.  This could be the start of further inquiry (e.g. possibly the cross referenced with the results from feature_selection.</p>
```{r Summary, warning=F, message=F}
```