---
title: "Wu_Kevin_Homework_1_HS650_Fall_2021.Rmd"
subtitle: "<h2><u>Data Science and Predictive Analytics (HS650), Fall 2021</u></h2>"
author: "<h3>Kevin Wu</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**Homework_1_HS650_Fall_2021.Rmd**.

 * HW #1
 * Fall 2021, DSPA (HS650)
 * Name: Kevin Wu
 * SID: ####0012 (last 4 digits only)
 * UMich E-mail: kevinkwu@umich.edu
 * I certify that the following paper represents my own independent work and conforms with the guidelines of academic honesty described in the UMich student handbook.
 * Remember that students are allowed and encouraged to discuss, on a conceptual level, the problems with your class mates, however, this can not involve the exchange of actual code, printouts, solutions, e-mails or other explicit electronic or paper handouts.


# HW Problem 1.1 
Problem 1.1 (Long-to-Wide Data format translation):

We demonstrated the wide-to-long conversion in lecture. Now, let’s explore long-to-wide. Load in the SOCR Pediatric Normal and Schizophrenia Neuroimaging Study Dataset (Links to an external site.). It's in a wide format where 56 different regional cortical and subcortical volumes contain information we want to represent in long format.

```{r Problem 1.1, warning=F, message=F}
library(XML)
library(reshape)
library(reshape2)
```

## HW Problem 1.1a
```{r Problem 1.1.a, warning=F, message=F}
SOCR_Data <- readHTMLTable("http://wiki.stat.ucla.edu/socr/index.php?title=SOCR_Data_Oct2009_ID_NI", header=TRUE, trim=TRUE, colClasses=rep("numeric",66)) 
df <- data.frame(SOCR_Data[[2]])

```


## HW Problem 1.1b 
Convert the original from wide-to-long
```{r Problem 1.1.b, warning=F, message=F}
dfLong <- melt(df, id.vars = c("X.Subject", "Age", "DX", "Sex"))

```


## HW Problem 1.1c
Then back to wide format.
```{r Problem 1.1.c, warning=F, message=F}
dfWide <- cast(dfLong)

```
 


# HW Problem 1.2 
Problem 1.2 (Data stratification):

Use the same Schizophrenia Neuroimaging Study Dataset (Links to an external site.) and complete the following data-manipulation steps in R:
```{r Problem2, warning=F, message=F}
library(plotly)

```

## HW Problem 1.2.a
Extract the first 10 subjects
```{r Problem2a, warning=F, message=F}
head(df,10)

```

## HW Problem 1.2.b
Find the cases for which L_caudate < 160.
```
subset(df, df$L_caudate < 160)
```

## HW Problem 1.2.c
Sort the subjects based on L_caudate values in descending.
```{r Problem2c, warning=F, message=F}
df[order(-df$L_caudate),]

```


## HW Problem 1.2.d
...and ascending order.
```{r Problem2d, warning=F, message=F}
df[order(df$L_caudate),]
```


## HW Problem 1.2.e
Generate frequency and probability tables for Age
```{r Problem2e, warning=F, message=F}
hist(df$Age, freq=F, breaks=max(df$Age) - min(df$Age), main="Density of Children's Ages Sampled", xlab="Age")
hist(df$Age, freq=T, breaks=max(df$Age) - min(df$Age), main="Frequency of Children's Ages Sampled", xlab="Age")
```


## HW Problem 1.2.f
Generate frequency and probability tables for IQ
```{r Problem2f, warning=F, message=F}
hist(df$FS_IQ, freq=F, breaks=max(df$FS_IQ) - min(df$FS_IQ), main="Density of Children's IQ's Sampled", xlab="IQ")
hist(df$FS_IQ, freq=T, breaks=max(df$FS_IQ) - min(df$FS_IQ), main="Frequency of Children's IQ's Sampled", xlab="IQ")
```

## HW Problem 1.2.g
Generate frequency and probability tables for Sex
```{r Problem2g, warning=F, message=F}
plot_ly(x=c("Male", "Female"), y=c(nrow(subset(df, df$Sex==1)), nrow(subset(df, df$Sex==2))))  %>%
  layout(title="Frequency of Children's Sex Sampled",
         xaxis=list(title="Sex", autotypenumbers="strict"),
         yaxis=list(title="Frequency"))

plot_ly(x=c("Male", "Female"), y=c(nrow(subset(df, df$Sex==1))/length(df), nrow(subset(df, df$Sex==2))/length(df))) %>%
    layout(title="Density of Children's Sex Sampled",
         xaxis=list(title="Sex", autotypenumbers="strict"),
         yaxis=list(title="Density"))
```

## HW Problem 1.2.h
Compute the mean Age and the correlation between Age and FS_IQ.
```{r Problem2h, warning=F, message=F}
mean(df$Age)
cor(df$Age, df$FS_IQ)
```


## HW Problem 1.2.i
Plot Histogram and density of R_fusiform_gyrus
```{r Problem2i, warning=F, message=F}
hist(df$R_fusiform_gyrus, main="Histogram of Right Fusiform Gyrus", xlab="Volumes in mm^3")
```

Draw scatterplot L_fusiform_gyrus and L_insular_cortex.
## HW Problem 1.2.j
```{r Problem2j, warning=F, message=F}
ggplot(df, aes(x=L_insular_cortex, y=L_fusiform_gyrus, )) + geom_point()

```
 
# HW Problem 1.3 

Problem 1.3 (Simulation)
Generate 1,000 standard normal variables and another 1,200 student t distributed random variables with df=20.  Generate a quantile-quantile (Q-Q) probability plot of the two samples.  Then, compare it with qqnorm() or plot_ly()  of student t simulation and interpret the findings.


## HW Problem 1.3a.
Generate 1,000 standard normal variables 
```{r name13a, warning=F, message=F}
valuesNorm <- rnorm(1000)
```

## HW Problem 1.3b.
and another 1,200 student t distributed random variables with df=20.
```{r name13b, warning=F, message=F}
valuesT <- rt(1200, 20)
```

## HW Problem 1.3c.
Generate a quantile-quantile (Q-Q) probability plot of the two samples.
```{r name13c, warning=F, message=F}
qValuesNorm <- quantile(scale(valuesNorm), probs=seq(from=0.01, to=0.99, by=0.01))
qValuesT <- quantile(scale(valuesT), probs=seq(from=0.01, to=0.99, by=0.01))
qq.df.norm_norm <- data.frame(qValuesNorm, qValuesT)

qq.df.norm_norm %>%
  plot_ly(x=~qValuesNorm) %>%
    add_markers(y=~qValuesT, name="Normally Distributed vs T-Distributed(df=20) Data") %>%
    add_lines(x=~qValuesNorm, y=~qValuesNorm, mode="line", name="Theoretical Model", line=list(width=2)) %>%
  layout(
    title="Normal Q-Q Plot", legend=list(orientation='h'))
```

## HW Problem 1.3d.
Then, compare it with qqnorm() or plot_ly()  of student t simulation and interpret the findings.
```{r name13d, warning=F, message=F}
set.seed(123)
qqnorm(valuesNorm)
qqline(valuesT)
```


# HW Problem 1.4 

Problem 1.4 (Define an R median function)
Define a new function  myMedian() that computes a sample median and compare your function against the standard R median() function using the simulation data you generate in the last question (1.3). Did you cover all possible situations for the input data?

<p>I believe I did cover all possible situations as I relied heavily upon R built in math functions.  Don't reinvent the wheel is my motto.
Utilizing R's built in rules of coersion, it's easy to assume that using their built in functions will also handle different data types "correctly" when using them with mathematical functions</p>
<p>
The only questionable things I do are
  <ol>
    <li>numvalues %% 2: is a very specific test for integers.  Dividing any object by 2 to test whether an integer is odd or even is a risky operation that may have many failure states, or worse unintended side-effects.  However, based upon R's definition of length, it does return a <i>non-negative integer of length 1</i>.  So this operation, based upon R's implementation is safe.  By definition of parity, this test does test whether the number of values is odd or even</li>
    <li>Setting the median index to round(numValues / 2): in the else case, numValues can be either 0, 1 or greater, but odd.  In the case for numValues is 0, then that is like doing a median of NA, which should return NA as it does.<br/>
    Note: <i>for rounding off a 5, the IEC 60559 standard (see also ‘IEEE 754’) is expected to be used, ‘go to the even digit’. Therefore round(0.5) is 0 and round(-1.5) is -2.</i><br/>
        &#x2022; In the case for numValues is 1, round returns the value of 0, which is desired.<br/>
        &#x2022; In the case for numValues is odd, round returns the next higher even number, which is also desired.
    </li>
    <li>I use mean because object at given index may be a complex object / data structure.  Leveraging the mean from the R library, the only non-trivial case is if na_REAL is returned, which is tested for and treated as if doing a median of NA</li>
  </ol>
</p>

```{r name4, warning=F, message=F}
myMedian <- function(values) {
  if(is.na(values))
    return(NA)
  else {
    values <- sort(values, na.last=NA)
    numValues <- length(values)
    # if even
    if((numValues %% 2) == 0) {
      mIndex1 <- numValues / 2
      mIndex2 <- mIndex1 + 1
      mnAtIndex1 <- mean(values[mIndex1])
      mnAtIndex2 <- mean(values[mIndex2])
      if(!is.na(mnAtIndex1) && !is.na(mnAtIndex2))
        return((mnAtIndex1+mnAtIndex2)/2)
      else
        return(NA)
    }
    else {
      if(numValues==0)
        return(NA)
      else {
        medianIndex <- round(numValues/2)
        return(values[medianIndex])        
      }
    }
  }
}
median(valuesNorm)
median(valuesT)
myMedian(valuesNorm)
myMedian(valuesT)
```
