---
title: "Wu_Kevin_Homework_2_HS650_Fall_2021.Rmd"
subtitle: "<h2><u>Data Science and Predictive Analytics (HS650), Fall 2021</u></h2>"
author: "<h3>Kevin Wu</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**Homework_2_HS650_Fall_2021.Rmd**.

 * HW #2
 * Fall 2021, DSPA (HS650)
 * Name: Kevin Wu
 * SID: ####0012 (last 4 digits only)
 * UMich E-mail: kevinkwu@umich.edu
 * I certify that the following paper represents my own independent work and conforms with the guidelines of academic honesty described in the UMich student handbook.
 * Remember that students are allowed and encouraged to discuss, on a conceptual level, the problems with your class mates, however, this can not involve the exchange of actual code, printouts, solutions, e-mails or other explicit electronic or paper handouts.


# HW Problem 2.1 
Load the following two datasets separately, generate summary statistics for all features, plot some of the features using plot_ly() histograms, box plots, density plots, etc., as appropriate, and save the summaries locally as Text files.
```{r Problem 2.1, warning=F, message=F}
library(reshape)
library(reshape2)
library(DT)
library(plotly)
library(rvest)
```

## Load ALS Data
```{r Load ALS Data, warning=F, message=F}
pathname_HW_2 <- "C:/Users/Kevin/Downloads/HS650 - Data Science and Predictive Analytics/Homework 2/"
pathname_ALS_small <- paste(pathname_HW_2, "ALS_TestingData_78.csv", sep="")
pathname_ALS_big <- paste(pathname_HW_2, "ALS_TrainingData_2223.csv", sep="")
dfALS <- read.csv(pathname_ALS_big)
datatable(dfALS)

```
## ALS Summary Statistics
```{r ALS Summary Statistics, warning=F, message=F}
textSummaryALS <- summary(dfALS)
textSummaryALS

```
## Write ALS Summary Statistics to Text File
```{r ALS Text File, warning=F, message=F}
fileALS <- file(paste(pathname_HW_2, "ALS_Summary.txt", sep=""))
capture.output(textSummaryALS, file=fileALS)

```
## Plot Features of ALS Statistics
```{r Plot Features of ALS Statistics, warning=F, message=F}
plot_ly(x = ~dfALS$hands_median, type="box", name="hands_median") %>%
  add_trace(x = ~dfALS$leg_median, name="leg_median") %>%
    layout(title="Median Leg and Hands", 
           xaxis=list(title="Medians"))

plot_ly(x = ~dfALS$Age_mean, y = ~dfALS$hands_median) %>%
    layout(title="Median Hands Given Age",
           xaxis=list(title="Median Age", yaxis=list(title="Median Hands")))

```
## Load SOCR Knee Pain Data
```{r Load SOCR Knee Pain Data Data, warning=F, message=F}
urlSOCR <- read_html("https://wiki.socr.umich.edu/index.php/SOCR_Data_KneePainData_041409")
dfSOCR <- html_table(html_nodes(urlSOCR, "table")[[2]])
datatable(dfSOCR)

```
## SOCR Knee Pain Summary Statistics
```{r SOCR Knee Pain Statistics, warning=F, message=F}
textSummarySOCR <- summary(dfSOCR)
textSummarySOCR

```
## Write SOCR Knee Pain Summary Statistics to Text File
```{r Knee Pain Text File, warning=F, message=F}
fileSOCR <- file(paste(pathname_HW_2, "SOCR_Summary.txt", sep=""))
capture.output(textSummarySOCR, file=fileSOCR)

```
## Plot Features of SOCR Knee Pain Statistics
```{r Plot Features of SOCR Knee Pain, warning=F, message=F}
plot_ly(x = ~dfSOCR$x, y = ~dfSOCR$Y, z= ~dfSOCR$View, type="scatter3d", mode="markers", size=0.5) %>%
    layout(yaxis=list(title='y location'), zaxis=list(title='View'), title='SOCR Knee Pain Summary', xaxis=list(title='x location'))

```
# HW Problem 2.2
- Use ALS case-study data and SOCR Knee Pain Data (Links to an external site.) to explore some bivariate relations (e.g. bivariate plot, correlation, table crosstable etc.)
- Use 07_UMich_AnnArbor_MI_TempPrecipitation_HistData_1900_2015 data to show the relations between temperature and time. [Hint: use geom_line or geom_bar]. Some sample code is included below.
```
library(gmodels)
```
## Exploring ALS Bivariate Relations
```{r Exploring ALS Bivariate Relations, warning=F, message=F}

plot_ly(x = ~dfALS$ALSFRS_Total_median, y = ~dfALS$hands_median, type = "scatter", mode = "markers", name="Hands") %>%
  add_trace(y = ~dfALS$leg_median, name="Legs") %>%
  add_trace(y = ~dfALS$mouth_median, name="Mouth") %>%
  layout(yaxis=list(title='Medians'), title='Medians Given Median ALSFRS Score',
     xaxis=list(title='Median ALSFRS Score'))

plot_ly(x = ~dfALS$ALSFRS_Total_median, y = ~dfALS$Blood.Urea.Nitrogen..BUN._median, type = "scatter", mode = "markers", name="Median BUN Age") %>%
  layout(yaxis=list(title='Median BUN'), title='Median BUN Given Median ALSFRS Score',
     xaxis=list(title='Median ALSFRS Score'))
```

## Exploring Temperature and Time Bivariate Relations
```{r Exploring Temperature and Time Bivariate Relations, warning=F, message=F}

  dfTemp <- as.data.frame(read.csv("https://umich.instructure.com/files/706163/download?download_frd=1", header=T, na.strings=c("", ".", "NA", "NR")))

  summary(dfTemp)

  dfLongTemp <- melt(dfTemp, id.vars = c("Year"))
  summary(dfLongTemp)
  plot_ly(x = ~dfLongTemp$variable, y = ~dfLongTemp$value, type = "box") %>%
    layout(yaxis=list(title='Temperature'), 
            xaxis=list(title='Month'), 
            title='Ann Arbor\'s Temperature in Faranheit by Month from 1900 to 2015')
  
```

# HW Problem 2.3
Introduce (artificially) some missing data in the Knee Pain dataset (Links to an external site.), impute the missing values and examine the differences between the original, incomplete, and imputed datasets.
```{r Imputations, warning=F, message=F}

library(Amelia)

create.missing <- function (data, pct.mis = 10) 
{
    n <- nrow(data)
    J <- ncol(data)
    if (length(pct.mis) == 1) {
        if(pct.mis>= 0 & pct.mis <=100) {
            n.mis <- rep((n * (pct.mis/100)), J)
        }
        else {
          warning("Percent missing values should be an integer between 0 and 100! Exiting"); break
        }
      }
    else {
        if (length(pct.mis) < J) 
            stop("The length of the missing-vector is not equal to the number of columns in the data! Exiting!")
            n.mis <- n * (pct.mis/100)
    }
    for (i in 1:ncol(data)) {
        if (n.mis[i] == 0) { # if column has no missing do nothing.
            data[, i] <- data[, i]
        }
        else {
            data[sample(1:n, n.mis[i], replace = FALSE), i] <- NA
              # For each given column (i), sample the row indices (1:n), 
              # a number of indices to replace as "missing", n.mis[i], "NA", 
          # without replacement
        }
    }
    return(as.data.frame(data))
}

#dfMissingSOCR <-# create.missing(cbind.data.frame(dfSOCR$x, dfSOCR$Y), pct.mis = 30)
#dfMissingSOCR$View <- dfSOCR$View
dfMissingSOCR <- create.missing(dfSOCR, pct.mis = 20)
datatable(dfMissingSOCR)

dfImputatedSOCR <- amelia(dfMissingSOCR,noms=c("View"))
dfImputatedSOCR <- na.omit(dfImputatedSOCR$imputations[[5]])
datatable(dfImputatedSOCR)
```

## Knee Pain Dataset Comparisons
Introduce (artificially) some missing data in the Knee Pain dataset (Links to an external site.), impute the missing values and examine the differences between the original, incomplete, and imputed datasets.
```{r Comparisons, warning=F, message=F}
plot_ly(x = ~dfSOCR$x, y = ~dfSOCR$Y, z= ~dfSOCR$View, type="scatter3d", mode="markers", size=0.5) %>%
    layout(yaxis=list(title='y location'), zaxis=list(title='View'), title='ORIGINAL Knee Pain Summary', xaxis=list(title='x location'))
plot_ly(x = ~dfMissingSOCR$x, y = ~dfMissingSOCR$Y, z= ~dfMissingSOCR$View, type="scatter3d", mode="markers", size=0.5) %>%
    layout(yaxis=list(title='y location'), zaxis=list(title='View'), title='MISSING Knee Pain Summary', xaxis=list(title='x location'))
plot_ly(x = ~dfImputatedSOCR$x, y = ~dfImputatedSOCR$Y, z= ~dfImputatedSOCR$View, type="scatter3d", mode="markers", size=0.5) %>%
    layout(yaxis=list(title='y location'), zaxis=list(title='View'), title='IMPUTED Knee Pain Summary', xaxis=list(title='x location'))

```
# HW Problem 2.4
Generate a surface plot for the (RF) Knee Pain data illustrating the 2D distribution of locations of the patient reported knee pain (use plot_ly and kernel density estimation).
```{r Surface Plot, warning=F, message=F}

dfSOCRRF <- head(subset(dfSOCR, dfSOCR$View == "RF", select=c("x","Y")), 100)
listKneeRF <- list()
listKneeRF$x <- dfSOCRRF$x
listKneeRF$y <- dfSOCRRF$Y
listKneeRF$z <- log(t(dfSOCRRF$x) %x% dfSOCRRF$Y)

with(listKneeRF, plot_ly(x=x, y=y, z=z, type="surface"))
```

# HW Problem 2.5
Rebalance the groups of ALS (training data) patients according  Age>50  vs. Age≤50, based on the synthetic minority oversampling (SMOTE) (Links to an external site.) to ensure approximately equal cohort sizes.
```{r ALS SMOTE, warning=F, message=F}

library(unbalanced)

#binarialize age into a separate column
dfALS$Less50 <- ifelse(dfALS$Age_mean <= 50, 1, 0)
dfALS$Less50 <- as.factor(dfALS$Less50)

#count
print(length(which(dfALS$Less50==1)))
print(length(which(dfALS$Less50==0)))

#rebalance the data using the SMOTE algorithm
input <- dfALS
output <- as.factor(dfALS$Less50)
dataOutput <- ubBalance(X=input, Y=output, type="ubSMOTE", percOver=100, verbose=TRUE)
table(dataOutput$Y)

balancedALS <- cbind(dataOutput$X, dataOutput$Y)
colnames(balancedALS) <- c(colnames(input), "Less50")

```
## Compare Raw Data and Rebalanced Data Using QQ Plot
```{r QQ Plot Check, warning=F, message=F}
QQ <- qqplot(input[, "ALSFRS_Total_median"], balancedALS [, "ALSFRS_Total_median"], plot.it=F)

plot_ly(x=~QQ$x, y = ~QQ$y, type="scatter", mode="markers", showlegend=F) %>%
  add_lines(x=c(0,48), y=c(0,48), showlegend=F) %>%
  layout(title="QQ-Plot Original vs. Rebalanced Data for ALSFRS_Total_median", xaxis=list(title="Original Data"),
         yaxis=list(title="Rebalanced Data"))

```
## Compare Raw Data and Rebalanced Data Using Wilcoxon Check
```{r Wilconxon Check, warning=F, message=F}
alpha05 <- 0.05
test.results.bin <- NULL        # binarized/dichotomized p-values
test.results.raw <- NULL        # raw p-values

for (i in 1:(ncol(balancedALS)-1)) {
  test.results.raw[i] <- wilcox.test(as.numeric(input[, i]), as.numeric(balancedALS [, i]))$p.value
  test.results.bin[i] <- ifelse(test.results.raw[i] > alpha05, 1, 0)
}
# zeros (0) are significant independent between-group T-test differences, ones (1) are insignificant
test.results.corr <- stats::p.adjust(test.results.raw, method = "fdr", n = length(test.results.raw)) 


plot_ly(x=~test.results.raw, y = ~test.results.corr, type="scatter", mode="markers", showlegend=F) %>%
  add_lines(x=c(0,1), y=c(0,1), showlegend=F) %>%
  layout(title="Wilcoxon test results - Original vs. Rebalanced Data", xaxis=list(title="Original"),
         yaxis=list(title="Rebalanced"))