---
title: "Homework 3"
subtitle: "<h2><u>Data Science and Predictive Analytics (HS650), Fall 2021</u></h2>"
author: "<h3>Kevin Wu</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**Homework_3_HS650_Fall_2021.Rmd**.

 * HW #3
 * Fall 2021, DSPA (HS650)
 * Name: Kevin Wu
 * SID: ####0012 (last 4 digits only)
 * UMich E-mail: kevinkwu@umich.edu
 * I certify that the following paper represents my own independent work and conforms with the guidelines of academic honesty described in the UMich student handbook.
 * Remember that students are allowed and encouraged to discuss, on a conceptual level, the problems with your class mates, however, this can not involve the exchange of actual code, printouts, solutions, e-mails or other explicit electronic or paper handouts.


# HW Problem 3.1 
For each of the probability distributions: Weibull(1,5), Uniform(1,10), Student's T (df=1), Cauchy, Negative Binomial(10, 0.5), Chi-Square (df=10), Poisson (5), 

1) Generate the following plots functions:
  <ul>
    <li>Density</li>
    <li>CDF</li>
    <li>quantile (inverse-CDF)</li>
  </ul>
2) Report the first 4 moments (mean, variance, skewness, kurtosis)
3) Complete the discrete probability distributions table below. The cell values in the table represent the values of the quantile function for the corresponding p-value (column) and distribution (row).

```{r Problem 3.1, warning=F, message=F}
library(reshape)
library(reshape2)
library(DT)
library(plotly)
library(rvest)
```

## 
```{r Load ALS Data, warning=F, message=F}

```