---
title: "Homework 3"
subtitle: "<h2><u>Data Science and Predictive Analytics (HS650), Fall 2021</u></h2>"
author: "<h3>Kevin Wu</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**Homework_3_HS650_Fall_2021.Rmd**.

 * HW #3
 * Fall 2021, DSPA (HS650)
 * Name: Kevin Wu
 * SID: ####0012 (last 4 digits only)
 * UMich E-mail: kevinkwu@umich.edu
 * I certify that the following paper represents my own independent work and conforms with the guidelines of academic honesty described in the UMich student handbook.
 * Remember that students are allowed and encouraged to discuss, on a conceptual level, the problems with your class mates, however, this can not involve the exchange of actual code, printouts, solutions, e-mails or other explicit electronic or paper handouts.

```{r Libararies, warning=F, message=F}
library(reshape)
library(reshape2)
library(DT)
library(plotly)
library(rvest)
library(moments)
library(factoextra)
library(Rtsne)
```

# HW Problem 3.1 
For each of the probability distributions: , Uniform(1,10), Student's T (df=1), Cauchy, Negative Binomial(10, 0.5), Chi-Square (df=10), Poisson (5), 
1) Generate the following plots functions:
  <ul>
    <li>Density</li>
    <li>CDF</li>
    <li>quantile (inverse-CDF)</li>
  </ul>
2) Report the first 4 moments (mean, variance, skewness, kurtosis)
3) Complete the discrete probability distributions table below. The cell values in the table represent the values of the quantile function for the corresponding p-value (column) and distribution (row).

## Weibull's (1,5) Distribution
```{r Weibull, warning=F, message=F}
i <- seq(0, 25, 0.1)
p <- seq(0.001, 0.999, 0.001)
dfWeib <- data.frame(x=i, density=dweibull(i, 1, 5), distribution=pweibull(i, 1, 5))
qDFWeib <- data.frame(x=p, quantile=qweibull(p, 1, 5))

plot_ly(x= ~dfWeib$x, y= ~dfWeib$density, type="scatter", mode="lines", name="PDF") %>%
  layout(title="Weibull's (1,5) Density Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~dfWeib$x, y= ~dfWeib$distribution, type="scatter", mode="lines", name="CDF") %>%
  layout(title="Weibull's (1,5) Distribution Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qDFWeib$x, y= ~qDFWeib$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Weibull's (1,5) Quantile Function", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```

## Uniform (1,10) Distribution
```{r Uniform, warning=F, message=F}
i <- seq(-45, 45, 0.1)
p <- seq(0.001, 0.999, 0.001)
dfNorm <- data.frame(x=i, density=dnorm(i, 1, 10), distribution=pnorm(i, 1, 10))
qDFNorm <- data.frame(x=p, quantile=qnorm(p, 1, 10))

plot_ly(x= ~dfNorm$x, y= ~dfNorm$density, type="scatter", mode="lines", name="PDF") %>%
  layout(title="Uniform (1,10) Density Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~dfNorm$x, y= ~dfNorm$distribution, type="scatter", mode="lines", name="CDF") %>%
  layout(title="Uniform (1,10) Distribution Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qDFNorm$x, y= ~qDFNorm$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Uniform (1,10) Quantile Function", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```

## Student's T (df=1) Distribution
```{r Student T, warning=F, message=F}
i <- seq(-10, 10, 0.1)
p <- seq(0.001, 0.999, 0.001)
dfT <- data.frame(x=i, density=dt(i, 1), distribution=pt(i, 1))
qDFT <- data.frame(x=p, quantile=qt(p, 1))

plot_ly(x= ~dfT$x, y= ~dfT$density, type="scatter", mode="lines", name="PDF") %>%
  layout(title="Student's T (df=1) Density Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~dfT$x, y= ~dfT$distribution, type="scatter", mode="lines", name="CDF") %>%
  layout(title="Student's T (df=1) Distribution Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qDFT$x, y= ~qDFT$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Student's T (df=1) Quantile Function", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```

## Cauchy Distribution
```{r Cauchy, warning=F, message=F}
i <- seq(-10, 10, 0.1)
p <- seq(0.001, 0.999, 0.001)
dfCauchy <- data.frame(x=i, density=dcauchy(i), distribution=pcauchy(i))
qDFCauchy <- data.frame(x=p, quantile=qcauchy(p))

plot_ly(x= ~dfCauchy$x, y= ~dfCauchy$density, type="scatter", mode="lines", name="PDF") %>%
  layout(title="Cauchy Density Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~dfCauchy$x, y= ~dfCauchy$distribution, type="scatter", mode="lines", name="CDF") %>%
  layout(title="Cauchy Distribution Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qDFCauchy$x, y= ~qDFCauchy$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Cauchy Quantile Function", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```

## Negative Binomial (10, 0.5) Distribution
```{r Negative Binomial, warning=F, message=F}
i <- seq(0, 20, 1)
p <- seq(0.001, 0.999, 0.001)
dfNBinom <- data.frame(x=i, density=dnbinom(i, 10, 0.5), distribution=pnbinom(i, 10, 0.5))
qDFNBinom <- data.frame(x=p, quantile=qnbinom(p, 10, 0.5))

plot_ly(x= ~dfNBinom$x, y= ~dfNBinom$density, type="scatter", mode="lines", name="PDF") %>%
  layout(title="Negative Binomial (10, 0.5) Density Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~dfNBinom$x, y= ~dfNBinom$distribution, type="scatter", mode="lines", name="CDF") %>%
  layout(title="Negative Binomial (10, 0.5) Distribution Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qDFNBinom$x, y= ~qDFNBinom$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Negative Binomial (10, 0.5) Qunatile Function", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```


## Chi-Square (df=10) Distribution
```{r Chi-Square Distribution, warning=F, message=F}
i <- seq(0, 20, 1)
p <- seq(0.001, 0.999, 0.001)
dfChisq <- data.frame(x=i, density=dchisq(i, 10), distribution=pchisq(i, 10))
qDFChisq <- data.frame(x=p, quantile=qchisq(p, 10))

plot_ly(x= ~dfChisq$x, y= ~dfChisq$density, type="scatter", mode="lines", name="PDF") %>%
  layout(title="Chi-Square (df=10) Density Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~dfChisq$x, y= ~dfChisq$distribution, type="scatter", mode="lines", name="CDF") %>%
  layout(title="Chi-Square (df=10) Distribution Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qDFChisq$x, y= ~qDFChisq$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Chi-Square (df=10) Quantile Function", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```

## Poisson (5) Distribution
```{r Poisson (5) Distribution, warning=F, message=F}
i <- seq(0, 20, 1)
p <- seq(0.001, 0.999, 0.001)
dfPoisson <- data.frame(x=i, density=dpois(i, 5), distribution=ppois(i, 5))
qDFPoisson <- data.frame(x=p, quantile=qpois(p, 5))

plot_ly(x= ~dfPoisson$x, y= ~dfPoisson$density, type="scatter", mode="lines", name="PDF") %>%
  layout(title="Poisson (5) Density Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~dfPoisson$x, y= ~dfPoisson$distribution, type="scatter", mode="lines", name="CDF") %>%
  layout(title="Poisson (5) Distribution Function", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qDFPoisson$x, y= ~qDFPoisson$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Poisson (5) Quantile Function", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```

## First 4 Moments: Mean, Variance, Skewness, Kurtosis
```{r Four Moments, warning=F, message=F}
listDens <- list()
listDens[[1]] <- as.list(dfWeib)["density"]
listDens[[2]] <- as.list(dfNorm)["density"]
listDens[[3]] <- as.list(dfT)["density"]
listDens[[4]] <- as.list(dfCauchy)["density"]
listDens[[5]] <- as.list(dfNBinom)["density"]
listDens[[6]] <- as.list(dfChisq)["density"]
listDens[[7]] <- as.list(dfPoisson)["density"]
dfResultsMoments <- data.frame(matrix(ncol=4, nrow=0))
for (l in listDens) {
  rDF <- c(sapply(l, mean), lapply(l, var), lapply(l, skewness), lapply(l, kurtosis))
  names(rDF) <- c("Mean", "Variance", "Skewness", "Kurtosis")
  dfResultsMoments <- rbind(dfResultsMoments, c(rDF$Mean, rDF$Variance, rDF$Skewness, rDF$Kurtosis))
}
colnames(dfResultsMoments) <- c("Mean","Variance","Skewness","Kurtosis")
rownames(dfResultsMoments) <- c("Weibull's(1,5)", "Uniform (1,10)", "Student's T (df=1)", "Cauchy", "Negative Binomial (10, 0.5)", "Chi-Square (df=10)", "Poisson (5)")
print(dfResultsMoments)
```

## Discrete Probability Distributions of the Quantile Functions
```{r Quantile Probabilities, warning=F, message=F}
dfProbabilities <- data.frame(matrix(ncol=0, nrow=7))
forSeq <- seq(0.1,0.9,0.1)
for (i in forSeq) {
  vecProb <- c(
    subset(qDFWeib, qDFWeib$x == i)$quantile,
    subset(qDFNorm, qDFNorm$x == i)$quantile,
    subset(qDFT, qDFT$x == i)$quantile,
    subset(qDFCauchy, qDFCauchy$x == i)$quantile,
    subset(qDFNBinom, qDFNBinom$x == i)$quantile,
    subset(qDFChisq, qDFChisq$x == i)$quantile,
    subset(qDFPoisson, qDFPoisson$x == i)$quantile
  )
  dfProbabilities[paste("P", i, "")] <- vecProb
}
rownames(dfProbabilities) <- c("Weibull's(1,5)", "Uniform (1,10)", "Student's T (df=1)", "Cauchy", "Negative Binomial (10, 0.5)", "Chi-Square (df=10)", "Poisson (5)")
print(dfProbabilities)
```

# HW Problem 3.2
Solve the following system of linear equations and validate your solution. Validate your solution.

6x + 3y - 3z + w = 2
7x + y + 2z + 2w = 5
5x + 3y - 3z + w = 3
-6x - 2y + 3z + 0w = 6
```{r Matrix Equations, warning=F, message=F}
A <- t(matrix(c(6,3,-3,1,7,1,2,2,5,3,-3,1,-6,-2,3,0),nrow = 4, ncol = 4))
b <- c(2,5,3,6)
xyzw <- solve(A, b)
xyzw
LHS <- A %*% xyzw
round(LHS-b, 6)

```

# HW Problem 3.3
Use PCA and t-SNE to analyze and interpret the monthly US Federal Reserve Monetary-Base Data (1959-2009)

## PCA Analysis
```{r PCA scree, warning=F, message=F}
urlUSFedRes <- read_html("http://wiki.stat.ucla.edu/socr/index.php/SOCR_Data_MonetaryBase1959_2009") 
dfFedResWithDates <- html_table(html_nodes(urlUSFedRes, "table")[[2]])
dfFedRes <- dfFedResWithDates[, -1]
savingsToCategory <- function(value) {
  if(value > 6000)
    category <- 4
  else if (value > 4000 && value <= 6000)
    category <- 3
  else if (value > 2000 && value <= 4000)
    category <- 2
  else
    category <- 1
  return(category)
}
catToColor <- function(category) {
  mapping = c("red","brown","blue","green")
  return(mapping[category])
}
train <- dfFedRes
train.labels <- sapply(dfFedRes$SAVINGSL, savingsToCategory)
train$label <- sapply(dfFedRes$SAVINGSL, savingsToCategory)
train$label <- as.factor(train$label)
train.labels.colors <- sapply(train$label, catToColor)
names(train.labels.colors) <- train$label


pca1 <- prcomp(as.matrix(dfFedRes), center=T, scale=T)
summary(pca1)

eigen<-get_eigenvalue(pca1)
eigen

#plot_ly(x = c(1:length(pca1$sdev)), y = (pca1$sdev)^2, name="Scree", type="bar") %>%
plot_ly(x = c(1:length(eigen$eigenvalue)), y = eigen$variance.percent, name="Scree", type="bar") %>%
  layout(title="Scree Plot", xaxis=list(title="PC's"), yaxis=list(title="Percentage"))
```
Based upon the PCA reduction model as shown by the the Scree bar chart, PC1 and PC2 accounts for the majority of the variations.


```{r PCA projections, warning=F, message=F}
scores <- pca1$x
loadings <- pca1$rotation
loadings
scaleLoad <- 100

p <- plot_ly() %>%
  add_trace(x=scores[,1], y=scores[,2], type="scatter", mode="markers", name=train$label,
    marker = list(color=train.labels.colors, opacity = 0.7), showLegend=F) %>%
  layout(title = "Dimensional Reduction Model of SOCR Data Monetary Base (1959-2009) from 7D-2D", 
         xaxis = list(title = rownames(loadings)[1]), yaxis = list(title = rownames(loadings)[2]))
p

modelFedRes <- princomp(dfFedRes, scale=T)
biplot(modelFedRes)
```


```{r PCA bootstrap, warning=F, message=F}
set.seed(12)
num_boot = 1000
bootstrap_it = function(i) {
  data_resample = dfFedRes[sample(1:nrow(dfFedRes), nrow(dfFedRes), replace=TRUE),] 
  p_resample = princomp(data_resample,cor = T) 
  return(sum(p_resample$sdev[1:3]^2)/sum(p_resample$sdev^2))
  }

pco = data.frame(per=sapply(1:num_boot, bootstrap_it)) 
quantile(pco$per, probs = c(0.025,0.975))  # specify 95-th % Confidence Interval

corpp = sum(pca1$sdev[1:3]^2)/sum(pca1$sdev^2)

plot_ly(x = pco$per, type = "histogram", name = "Data Histogram") %>% 
    layout(title='Histogram of a Bootstrap Simulation <br /> Percent of Data Variability Captured by first 3 PCs', 
           xaxis = list(title = "Percent of Variability"), yaxis = list(title = "Frequency Count"), bargap=0.1)
```

## t-SNE Analysis
Run the t-SNE, tracking the execution time (artificially reducing the sample-size to get reasonable calculation time)
```{r t-SNE analysis, warning=F, message=F}
execTime_tSNE <- system.time(tsne_FedRes <- Rtsne(train[,], dims = 2, perplexity=30, verbose=TRUE, max_iter = 1000)); execTime_tSNE

dfTSNE <- data.frame(tsne_FedRes$Y, train.labels.colors)
plot_ly(dfTSNE, x = ~X1, y = ~X2, mode = 'text') %>%
  add_text(text = names(train.labels.colors), textfont = list(color = dfTSNE$train.labels.colors)) %>%
  layout(title = "t-SNE 2D Embedding", xaxis = list(title = ""),  yaxis = list(title = ""))
```

# HW Problem 3.4
Lease Squares Estimation
Use the SOCR Knee Pain dataset (Links to an external site.), extract the RB = Right-Back locations (x,y), and fit in a linear model for vertical location (y) in terms of the horizontal location (x). Display the linear model on top of the scatter plot of the paired data.
```{r Least Squares, message=FALSE, warning=FALSE}
urlSOCR <- read_html("https://wiki.socr.umich.edu/index.php/SOCR_Data_KneePainData_041409")
dfSOCR <- html_table(html_nodes(urlSOCR, "table")[[2]])
dfSOCRRB <- subset(dfSOCR, dfSOCR$View=="RB")
fit <- lm(Y ~ x, data = dfSOCRRB)

plot_ly() %>%
  add_markers(x = dfSOCRRB$x, y = dfSOCRRB$Y, name="Pain Location") %>%
  add_lines(x = dfSOCRRB$x, y= fit$fitted, name="Least Squares") %>%
  layout(title = "SOCR Data Knee Pain Data Least Squares Estimation", 
         xaxis = list(title="X-Location"), yaxis=list(title="Y-Location"))