---
title: "Homework 3"
subtitle: "<h2><u>Data Science and Predictive Analytics (HS650), Fall 2021</u></h2>"
author: "<h3>Kevin Wu</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**Homework_3_HS650_Fall_2021.Rmd**.

 * HW #3
 * Fall 2021, DSPA (HS650)
 * Name: Kevin Wu
 * SID: ####0012 (last 4 digits only)
 * UMich E-mail: kevinkwu@umich.edu
 * I certify that the following paper represents my own independent work and conforms with the guidelines of academic honesty described in the UMich student handbook.
 * Remember that students are allowed and encouraged to discuss, on a conceptual level, the problems with your class mates, however, this can not involve the exchange of actual code, printouts, solutions, e-mails or other explicit electronic or paper handouts.


```{r Problem 3.1, warning=F, message=F}
library(reshape)
library(reshape2)
library(DT)
library(plotly)
library(rvest)
```
# HW Problem 3.1 
For each of the probability distributions: , Uniform(1,10), Student's T (df=1), Cauchy, Negative Binomial(10, 0.5), Chi-Square (df=10), Poisson (5), 

1) Generate the following plots functions:
  <ul>
    <li>Density</li>
    <li>CDF</li>
    <li>quantile (inverse-CDF)</li>
  </ul>
2) Report the first 4 moments (mean, variance, skewness, kurtosis)
3) Complete the discrete probability distributions table below. The cell values in the table represent the values of the quantile function for the corresponding p-value (column) and distribution (row).

## Weibull (1,5) Distribution
```{r Weibull, warning=F, message=F}
i <- seq(0, 10, 0.1)
p <- seq(0.001, 0.999, 0.001)
dfWeib <- data.frame(x=i, density=dweibull(i, 1, 5), distribution=pweibull(i, 1, 5))
qWeib <- data.frame(x=p, quantile=qweibull(p, 1, 5))

plot_ly(x= ~dfWeib$x, y= ~dfWeib$density, type="scatter", mode="lines", name="Density") %>%
  add_lines(x= ~dfWeib$x, y= ~dfWeib$distribution, name="Distribution") %>%
  layout(title="Weibull Curve", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qWeib$x, y= ~qWeib$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Weibull Quantile", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```

## Uniform (1,10) Distribution
```{r Uniform, warning=F, message=F}
i <- seq(-5, 5, 0.1)
p <- seq(0.001, 0.999, 0.001)
dfNorm <- data.frame(x=i, density=dnorm(i, 0, 1), distribution=pnorm(i, 0, 1))
qNorm <- data.frame(x=p, quantile=qnorm(p, 0, 1))

plot_ly(x= ~dfNorm$x, y= ~dfNorm$density, type="scatter", mode="lines", name="Density") %>%
  add_lines(x= ~dfNorm$x, y= ~dfNorm$distribution, name="Distribution") %>%
  layout(title="Uniform Curve", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qWeib$x, y= ~qWeib$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Uniform Distribution", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```
## Student's T (df=1)
```{r Student T, warning=F, message=F}
i <- seq(-5, 5, 0.1)
p <- seq(0.001, 0.999, 0.001)
dfT <- data.frame(x=i, density=dt(i, 1), distribution=pt(i, 1))
qT <- data.frame(x=p, quantile=qt(p, 1))

plot_ly(x= ~dfT$x, y= ~dfT$density, type="scatter", mode="lines", name="Density") %>%
  add_lines(x= ~dfT$x, y= ~dfT$distribution, name="Distribution") %>%
  layout(title="Student's T Curve", legend=list(orientation="h"), xaxis=list(title="X-Values"), yaxis=list(title="Y-Values"))
plot_ly(x= ~qT$x, y= ~qT$quantile, type="scatter", mode="lines", name="Quant") %>%
  layout(title="Student's T Distribution", legend=list(orientation="h"), xaxis=list(title="Probabilities"), yaxis=list(title="Values"))
```
