---
title: "Homework 4"
subtitle: "<h2><u>Data Science and Predictive Analytics (HS650), Fall 2021</u></h2>"
author: "<h3>Kevin Wu</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**Homework_4_HS650_Fall_2021.Rmd**.

 * HW #4
 * Fall 2021, DSPA (HS650)
 * Name: Kevin Wu
 * SID: ####0012 (last 4 digits only)
 * UMich E-mail: kevinkwu@umich.edu
 * I certify that the following paper represents my own independent work and conforms with the guidelines of academic honesty described in the UMich student handbook.
 * Remember that students are allowed and encouraged to discuss, on a conceptual level, the problems with your class mates, however, this can not involve the exchange of actual code, printouts, solutions, e-mails or other explicit electronic or paper handouts.

```{r Libararies, warning=F, message=F}
library(rvest)
library(tm)
library(wordcloud)
library(plotly)
library(e1071)
library(gmodels)
library(dplyr)
library(MASS)
library(C50)
library(caret)
library(rpart)
library(rpart.utils)
```

# HW Problem 4.1, Data Preparation, and Distributions
```{r 4.1, warning=F, message=F}
urlSOCR <- read_html("http://wiki.socr.umich.edu/index.php/SOCR_Data_2011_US_JobsRanking#2011_Ranking_of_the_200_most_common_Jobs_in_the_US")
socr_df <- html_table(html_nodes(urlSOCR, "table")[[1]])
jsd_df <- socr_df[-1]

plot_ly(jsd_df, x = ~Stress_Category, y = ~Hiring_Potential, type="box") %>%
  layout(title = "Hiring Potential Given Stress Category", xaxis = list(title = "Stress Category"), yaxis = list(title = "Hiring Potential"))
plot_ly(jsd_df, x = ~Hiring_Potential, type="histogram") %>%
  layout(title = "Hiring Potential Distribution", xaxis = list(title = "Hiring Potential"), bargap=0.1)
plot_ly(jsd_df, x = ~Stress_Category, type="histogram") %>%
  layout(title = "Stress Category Distribution", xaxis = list(title = "Stress Category"), bargap=0.1)
plot_ly(jsd_df, x = ~Stress_Category, y = ~`Average_Income(USD)`, type="box") %>%
  layout(title = "Average Income (USD) Given Stress Category", xaxis = list(title = "Stress Category"), yaxis = list(title = "Average Income (USD)"))
plot_ly(jsd_df, x = ~Physical_Demand, y = ~`Average_Income(USD)`, type="scatter") %>%
#  add_lines(x = ~Physical_Demand, y= ~lm( Physical_Demand ~ `Average_Income(USD)`)) %>%
  layout(title = "Average Income (USD) Given Physical Demand", xaxis = list(title = "Physical Demand"), yaxis = list(title = "Average Income (USD)"))

jsd_df$Description <- gsub('_', ' ', jsd_df$Description)
jsd_corpus <- Corpus(VectorSource(jsd_df$Description))
corpus_clean <- tm_map(jsd_corpus, tolower)
corpus_clean <- tm_map(corpus_clean, removePunctuation)
corpus_clean <- tm_map(corpus_clean, stripWhitespace)
corpus_clean <- tm_map(corpus_clean, removeNumbers)
jsd_dtm <- DocumentTermMatrix(corpus_clean)

set.seed(12345)
subsetInterval <- sample(nrow(jsd_df), floor(nrow(jsd_df) * 0.9))  # 90% training + 10% testing
jsd_df_train <- jsd_df[subsetInterval, ]
jsd_df_test <- jsd_df[-subsetInterval, ]
jsd_dtm_train <- jsd_dtm[subsetInterval, ]
jsd_dtm_test <- jsd_dtm[-subsetInterval, ]
corpus_train <- corpus_clean[subsetInterval]
corpus_test <- corpus_clean[-subsetInterval]

#See Distribution of Stress Categories between Training vs Testing datasets
prop.table(table(jsd_df_train$Stress_Category))
prop.table(table(jsd_df_test$Stress_Category))

#Binari"ize" Stress label for training and testing datasets
jsd_df_train$Stress_High <- jsd_df_train$Stress_Category %in% c(3, 4, 5)
jsd_df_train$Stress_High <- factor(jsd_df_train$Stress_High, levels=c(F, T), labels = c("Low_Stress", "High_Stress"))
jsd_df_test$Stress_High <- jsd_df_test$Stress_Category %in% c(3, 4, 5)
jsd_df_test$Stress_High <- factor(jsd_df_test$Stress_High, levels=c(F, T), labels = c("Low_Stress", "High_Stress"))

prop.table(table(jsd_df_train$Stress_High))
prop.table(table(jsd_df_test$Stress_High))
```

## Word Cloud and Graphics Between Low and High Stress Categories
```{r Low and High Stress Categories, warning=F, message=F}
# separate out low vs high stress from training data for graphics
jsd_df_train_low <- subset(jsd_df_train, Stress_High=="Low_Stress")
jsd_df_train_high <- subset(jsd_df_train, Stress_High=="High_Stress")
corpus_train_low <- Corpus(VectorSource(jsd_df_train_low$Description))
corpus_train_low <- tm_map(corpus_train_low, tolower)
corpus_train_low <- tm_map(corpus_train_low, removePunctuation)
corpus_train_low <- tm_map(corpus_train_low, removeNumbers)
corpus_train_low <- tm_map(corpus_train_low, stripWhitespace)
corpus_train_low <- tm_map(corpus_train_low, removeWords, stopwords('english'))
corpus_train_high <- Corpus(VectorSource(jsd_df_train_high$Description))
corpus_train_high <- tm_map(corpus_train_high, tolower)
corpus_train_high <- tm_map(corpus_train_high, removeNumbers)
corpus_train_high <- tm_map(corpus_train_high, removePunctuation)
corpus_train_high <- tm_map(corpus_train_high, stripWhitespace)
corpus_train_high <- tm_map(corpus_train_high, removeWords, stopwords('english'))
jsd_dtm_train_low <- DocumentTermMatrix(corpus_train_low)
jsd_dtm_train_low <- removeSparseTerms(jsd_dtm_train_low, 0.98)
jsd_dtm_train_high <- DocumentTermMatrix(corpus_train_high)
jsd_dtm_train_high <- removeSparseTerms(jsd_dtm_train_high, 0.98)
# Low Stress Jobs
wordcloud(corpus_train_low, min.freq = 7, random.order = FALSE, colors = brewer.pal(4, "Dark2"))
wordSumLow <- colSums(as.matrix(jsd_dtm_train_low))
plot_ly(x = ~unname(wordSumLow), y = ~names(wordSumLow), type="bar", orientation="h") %>%
  layout(title="Common Word Descriptions Low Stress Jobs", xaxis = list(title = "Words"), yaxis = list(title = "Counts"), bargap="0.1")
# High Stress Jobs
wordcloud(corpus_train_high, min.freq = 3, random.order = FALSE, colors = brewer.pal(4, "Dark2"))
wordSumHigh <- colSums(as.matrix(jsd_dtm_train_high))
plot_ly(x = ~unname(wordSumHigh), y = ~names(wordSumHigh), type="bar", orientation="h") %>%
  layout(title="Common Word Descriptions High Stress Jobs", xaxis = list(title = "Words"), yaxis = list(title = "Counts"), bargap="0.1")
```

## Naive Bayes
```{r Naive Bayes, warning=F, message=F}
# Create Function to binary"ize" Yes/No whether word appears > 5 times in the DTM and Apply
convert_counts <- function(x) {
  x <- ifelse(x > 0, 1, 0)
  x <- factor(x, levels = c(0, 1), labels = c("No", "Yes"))
  return(x)
}

#create training dictionary object & respective DTMs
jsd_dict <- as.character(findFreqTerms(jsd_dtm_train, 5))
jsd_train <- DocumentTermMatrix(corpus_train, list(dictionary = jsd_dict))
jsd_test <- DocumentTermMatrix(corpus_test, list(dictionary = jsd_dict))
jsd_train <- apply(jsd_train, MARGIN = 2, convert_counts)
jsd_test <- apply(jsd_test, MARGIN = 2, convert_counts)

jsd_bayes_classifier <- naiveBayes(jsd_train, jsd_df_train$Stress_High)
jsd_bayes_presdicter <- predict(jsd_bayes_classifier, jsd_test)
bayes_ct <- CrossTable(jsd_bayes_presdicter, jsd_df_test$Stress_High)
bayes_ct

bayes_TN <- bayes_ct$prop.row[1, 1]
bayes_FP <- bayes_ct$prop.row[1, 2]
bayes_FN <- bayes_ct$prop.row[2, 1]
bayes_TP <- bayes_ct$prop.row[2, 2]

bayes_acc <- (bayes_TP + bayes_TN) / (bayes_TP + bayes_FP + bayes_FN + bayes_TN)
bayes_acc

bayes_sens <- bayes_TP / (bayes_TP + bayes_FN)
bayes_sens

bayes_spec <- bayes_TN / (bayes_TN + bayes_FP)
bayes_spec

```

## LDA
```{r LDA, warning=F, message=F}
binarizeFunction <- function(x) { ifelse(x=="Yes", 1,0) }
cat2Numeric <- function (dfInput) {
  df = as.data.frame(lapply( as.data.frame(dfInput), factor)) %>%
                       mutate_all(binarizeFunction)
   return(df)
}
lda_df_train = data.frame(cat2Numeric(jsd_train), Stress_High = as.numeric(jsd_df_train$Stress_High))
lda_df_test = data.frame(cat2Numeric(jsd_test), Stress_High = as.numeric(jsd_df_test$Stress_High))
#did not bother removing any VIF's multi collinearities

jsd_lda_classifier <- lda(data = lda_df_train, Stress_High~.)
jsd_lda_predict <- predict(jsd_lda_classifier, lda_df_test)
lda_ct <- CrossTable(jsd_lda_predict$class, lda_df_test$Stress_High)
lda_ct

lda_TN <- lda_ct$prop.row[1, 1]
lda_FP <- lda_ct$prop.row[1, 2]
lda_FN <- lda_ct$prop.row[2, 1]
lda_TP <- lda_ct$prop.row[2, 2]

lda_acc <- (lda_TP + lda_TN) / (lda_TP + lda_FP + lda_FN + lda_TN)
lda_acc

#this can't be right...
lda_sens <- lda_TP / (lda_TP + lda_FN)
lda_sens

lda_spec <- lda_TN / (lda_TN + lda_FP)
lda_spec

```

## C5.0
```{r C5.0, warning=F, message=F}
c50_df <- as.data.frame(jsd_train, stringsAsFactors = T)
jsd_tree_50 <- C5.0(x = c50_df, y = jsd_df_train$Stress_High, trials = 3)
jsd_tree_50
summary(jsd_tree_50)

jsd_tree_50_predict <- predict(jsd_tree_50, jsd_test)
confusionMatrix(table(jsd_tree_50_predict, jsd_df_test$Stress_High))

## UNABLE to plot - Generates following error
# Error in str2lang(x) : <text>:1:28: unexpected '+'
# 1: y ~ and + commercial + for +
#  plot(jsd_tree_50)

```

## Multiple Linear Regression
```{r Multiple Linear Regression, warning=F, message=F}

fit <- lm(formula = Overall_Score ~ `Average_Income(USD)` + Work_Environment + Stress_Level + Stress_Category + Physical_Demand + Hiring_Potential, jsd_df)
summary(fit)

plot_ly(x=fit$fitted.values, y=fit$residuals, type="scatter", mode="markers") %>%
  layout(title="LM: Fitted-values vs. Model-Residuals",
         xaxis=list(title="Fitted"), 
         yaxis = list(title="Residuals"))
step(fit,direction = "backward")
