---
title: "Homework 6"
subtitle: "<h2><u>Data Science and Predictive Analytics (HS650), Fall 2021</u></h2>"
author: "<h3>Kevin Wu</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
tags: [DSPA, SOCR, MIDAS, Big Data, Predictive Analytics] 
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    self_contained: yes
---
**Homework_6_HS650_Fall_2021.Rmd**.

 * HW #6
 * Fall 2021, DSPA (HS650)
 * Name: Kevin Wu
 * SID: ####0012 (last 4 digits only)
 * UMich E-mail: kevinkwu@umich.edu
 * I certify that the following paper represents my own independent work and conforms with the guidelines of academic honesty described in the UMich student handbook.
 * Remember that students are allowed and encouraged to discuss, on a conceptual level, the problems with your class mates, however, this can not involve the exchange of actual code, printouts, solutions, e-mails or other explicit electronic or paper handouts.

```{r Libararies, warning=F, message=F}
library(plotly)
library(igraph)
library(networkD3)

library(DT)
library(Boruta)
library(caret)
library(randomForest)
library(mlbench)
```

# HW Problem 6.1 (Network data visualization)
Use the "Les Miserables (Links to an external site.)" dataset (03_les miserablese_GraphData.txt)
Visualize the network of (novel) characters as an undirected graph.
Summarize the graph and explain the output.
Calculate the degree and the centrality of this graph.
Find some important nodes (corresponding to key novel characters).
Will the results change if we assume the graph is directed?

```{r 6.1, warning=F, message=F}
miserablese<-read.table("https://umich.instructure.com/courses/38100/files/330389/download?download_frd=1", sep="", header=F)
miserablese <- as.matrix(miserablese, ncol=2)
head(miserablese)

graph_m <- graph.edgelist(miserablese, directed = F)
# U - undirected graph
# N - named edge
# 77 - nodes
# 254 - edges or relationships
summary(graph_m)

df <- as_data_frame(graph_m, what = "edges")
# Javascript note indexing starts at zero, not 1, make an artificial index zero root
df1 <- rbind(c(0,1), df)   
# Use D3 to display graph
simpleNetwork(df1,fontSize = 10, zoom = T)

# The following nodes have the greatest number of edges, signalling their connectedness and perhaps importance
# Valjean		36
# Gavroche	22
# Marius		19
degree(graph_m)

# The following are the previous nodes' betweeness, suggesting for the most part that connectedness may indicate importance, aside from Myreal with the number of edges = 10, which is lower in comparison to other nodes.
# Valjean		1624.4688004
# Myreal		504.0000000
# Gavroche	470.5706319
# Marius		376.2925926
betweenness(graph_m)

# the results would change if the graph were directed in that having terminal nodes may have a high degree of edges / connections, but a lower degree of conntectedness / centrality as by definition, they are terminal nodes.
```

# HW Problem 6.2 (Feature Selection):
Use the 06_PPMI_ClassificationValidationData_Short dataset
Set ResearchGroup as class variable.
Delete irrelevant columns (e.g. X, FID_IID) and select only the PD and Control cohorts, as the ResearchGroup feature includes PD, Control and SWEDD (which are mild cases of PD)
Properly convert the variables types, as may be necessary
Compare alternative feature selection methods - Boruta, Random Feature Elimination (RFE), and stepwise feature selection
Summarize and visualize the results
Report and compare the variables selected by these methods. How much overlap is there in the selected salient features?

```{r 6.2, warning=F, message=F}
ppmi_df <- read.csv("https://umich.instructure.com/files/330400/download?download_frd=1")
filter <- c("X", "FID_IID", "UPDRS_Part_I_Summary_Score_Baseline", "UPDRS_Part_I_Summary_Score_Month_03", "UPDRS_Part_I_Summary_Score_Month_06", "UPDRS_Part_I_Summary_Score_Month_09", "UPDRS_Part_I_Summary_Score_Month_12", "UPDRS_Part_I_Summary_Score_Month_18", "UPDRS_Part_I_Summary_Score_Month_24", "UPDRS_Part_II_Patient_Questionnaire_Summary_Score_Baseline", "UPDRS_Part_II_Patient_Questionnaire_Summary_Score_Month_03", "UPDRS_Part_II_Patient_Questionnaire_Summary_Score_Month_06", "UPDRS_Part_II_Patient_Questionnaire_Summary_Score_Month_09", "UPDRS_Part_II_Patient_Questionnaire_Summary_Score_Month_12", "UPDRS_Part_II_Patient_Questionnaire_Summary_Score_Month_18", "UPDRS_Part_II_Patient_Questionnaire_Summary_Score_Month_24", "UPDRS_Part_III_Summary_Score_Baseline", "UPDRS_Part_III_Summary_Score_Month_03", "UPDRS_Part_III_Summary_Score_Month_06", "UPDRS_Part_III_Summary_Score_Month_09", "UPDRS_Part_III_Summary_Score_Month_12", "UPDRS_Part_III_Summary_Score_Month_18", "UPDRS_Part_III_Summary_Score_Month_24")
ppmi_df_filtered <- ppmi_df[ , !names(ppmi_df) %in% filter]
ppmi_df_filtered <- subset(ppmi_df_filtered, ResearchGroup != "SWEDD")
ppmi_df_filtered$ResearchGroup <- as.factor(ppmi_df_filtered$ResearchGroup)
ppmi_df_filtered <- as.data.frame(lapply(ppmi_df_filtered, as.integer))
datatable(ppmi_df_filtered)

set.seed(123)
boruta_model <- Boruta(ResearchGroup~., data = ppmi_df_filtered)
print(boruta_model)

df_long <- tidyr::gather(as.data.frame(boruta_model$ImpHistory), feature, measurement)
plot_ly(df_long, y = ~measurement, color = ~feature, type = "box") %>%
  layout(title="Box-and-whisker Plots Across All PPMI Features",
           xaxis = list(title="Features"),
           yaxis = list(title="Importance"),
           showlegend=F)

final_boruta_model <- TentativeRoughFix(boruta_model)

getConfirmedFormula(final_boruta_model)

print(final_boruta_model$finalDecision[final_boruta_model$finalDecision %in% c("Confirmed", "Tentative")])


set.seed(123)
control <- rfeControl(functions = rfFuncs, method = "cv", number=10)
rf.train <- rfe(ppmi_df_filtered[, !names(ppmi_df_filtered) %in% "ResearchGroup"], ppmi_df_filtered[, c("ResearchGroup")], sizes=c(10, 20, 30, 40), rfeControl=control)
rf.train
plot(rf.train, type="b")


predRFE <- predictors(rf.train)
predBoruta <- getSelectedAttributes(final_boruta_model, withTentative = F)

#  Boruta = RFE all 26 variables overlap
intersect(predBoruta, predRFE)


# Define a base model - intercept only
base.mod <- lm(ResearchGroup ~ 1, data = ppmi_df_filtered)
# Define the full model - including all predictors
all.mod <- lm(ResearchGroup ~ ., data = ppmi_df_filtered)
ols_step <- lm(ResearchGroup ~ ., data = ppmi_df_filtered)
ols_step <- step(base.mod, scope = list(lower = base.mod, upper = all.mod), direction = 'both', k=2, trace = F)
summary(ols_step) # ols_step

# get the shortlisted variable
stepwiseConfirmedVars <- names(unlist(ols_step[[1]]))
# remove the intercept
stepwiseConfirmedVars <- stepwiseConfirmedVars[!stepwiseConfirmedVars %in% "(Intercept)"]
print(stepwiseConfirmedVars)


# estimate variable importance
predStepwise <- varImp(ols_step, scale=FALSE)
# summarize importance
print(predStepwise)

# Boruta & OLS only 9 variables intersect
intersect(predBoruta, stepwiseConfirmedVars)

# RFE & OLS 12 variables intersect
intersect(predRFE, stepwiseConfirmedVars)

# Stepwise predicts the importance of the 14 variables.  Boruta and RFE variables overlap differently with stepwise.
# However, Top 5 RFE, Boruta, and OLS are all in agreement being the top 5 variables.  It is safe to say that these are good precitors for the outcome ResearchGroup variable.
# - X_Assessment_Non.Motor_Geriatric_Depression_Scale_GDS_Short_Summary_Score_Month_06
# - L_putamen_Volume
# - X_Assessment_Non.Motor_Epworth_Sleepiness_Scale_Summary_Score_Month_24
# - Age
# - chr17_rs199533_DP
```